---
title: "Movie Memory Experiment — Initial Data Visualizations"
author: ''
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    df_print: paged
subtitle: BRSM Project 2026
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = TRUE,
  message   = FALSE,
  warning   = FALSE,
  fig.align = "center",
  fig.width = 8,
  fig.height = 5
)
```

---

## Setup — Packages & Libraries

```{r libraries}
required_packages <- c(
  "ggplot2", "dplyr", "tidyr", "readr",
  "ggmosaic", "ggbeeswarm", "fmsb",
  "RColorBrewer", "scales", "stringr", "patchwork"
)

for (pkg in required_packages) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg, repos = "https://cloud.r-project.org")
  }
}

library(ggplot2); library(dplyr); library(tidyr); library(readr); library(ggbeeswarm); library(fmsb)
library(RColorBrewer); library(scales); library(stringr); library(patchwork)
```

---

## Load & Prepare Data

> Update file paths below to match where your CSVs are saved.

```{r load-data}
nat   <- read_csv("naturalmovies.csv")
abr   <- read_csv("abruptmovies.csv")
tl    <- read_csv("target_and_lures.csv")
demog <- read_csv("demographics.csv")   # update filename if different
```

```{r tidy-data}
# Condition comes from which file the video belongs to
nat <- nat %>% mutate(condition = "Natural (NB)")
abr <- abr %>% mutate(condition = "Abrupt (AB)")

# Combined stimulus dataset
movies <- bind_rows(nat, abr) %>%
  mutate(condition = factor(condition, levels = c("Natural (NB)", "Abrupt (AB)")))

# Derive frame type from target filename
tl <- tl %>%
  mutate(
    frame_type = case_when(
      str_detect(target_img, "_EM_") ~ "Event-Middle (EM)",
      str_detect(target_img, "_BB_") ~ "Before-Boundary (BB)",
      TRUE                           ~ "Other"
    )
  )

# Extract condition from Sub ID suffix (_NB or _AB)
demog <- demog %>%
  mutate(
    condition = case_when(
      str_detect(`Sub ID`, "_NB$|_NB_") ~ "Natural (NB)",
      str_detect(`Sub ID`, "_AB$|_AB_") ~ "Abrupt (AB)",
      TRUE ~ NA_character_
    ),
    condition = factor(condition, levels = c("Natural (NB)", "Abrupt (AB)"))
  )
```

```{r colour-palettes}
pal2  <- c("Natural (NB)" = "#4CAF50", "Abrupt (AB)" = "#E91E63")
pal_ft <- c(
  "Event-Middle (EM)"    = "#2196F3",
  "Before-Boundary (BB)" = "#FF9800",
  "Other"                = "#9E9E9E"
)
```

---

## Section 1 — Stimulus Properties

*These plots describe the videos shown during encoding and verify that the Natural and Abrupt conditions were matched on duration.*

---

### Fig 1 — Violin Plot: Video Duration by Condition

> **What it shows:** The full distribution of clip durations for Natural vs Abrupt videos.
> Since abrupt cuts interrupt videos 1–5 seconds before a boundary, Abrupt clips should
> be slightly shorter on average. This plot checks whether the duration matching procedure
> (described in the methods) was successful.

```{r violin-duration}
ggplot(movies, aes(x = condition, y = duration, fill = condition)) +
  geom_violin(trim = FALSE, alpha = 0.7, colour = "grey40") +
  geom_boxplot(width = 0.08, fill = "white", outlier.shape = NA) +
  scale_fill_manual(values = pal2) +
  labs(
    title    = "Fig 1 · Violin Plot: Video Duration by Condition",
    subtitle = "Natural vs Abrupt cut clips — checking duration matching",
    x = "Condition", y = "Duration (seconds)"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"), legend.position = "none")
```

---

### Fig 2 — Raindrop / Beeswarm Plot: Individual Video Durations

> **What it shows:** Every individual video as a single point, spread so they don't overlap.
> This reveals the actual data points behind the violin — useful for spotting outliers or
> clusters of unusually long/short clips in either condition.

```{r raindrop-duration}
ggplot(movies, aes(x = condition, y = duration, colour = condition)) +
  geom_beeswarm(size = 3, alpha = 0.75, cex = 2.5) +
  geom_boxplot(width = 0.2, alpha = 0, colour = "grey30", outlier.shape = NA) +
  scale_colour_manual(values = pal2) +
  labs(
    title    = "Fig 2 · Raindrop Plot: Individual Video Durations per Condition",
    subtitle = "Each point = one video clip",
    x = "Condition", y = "Duration (seconds)"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"), legend.position = "none")
```

---

### Fig 3 — Line Plot: Cumulative Duration Across Stimulus Order

> **What it shows:** How total encoding time accumulates across the stimulus list for each condition.
> The two lines should converge toward similar totals at the end — confirming that duration
> was matched between conditions as described in the methods.

```{r line-cumulative}
line_data <- movies %>%
  group_by(condition) %>%
  mutate(
    video_index  = row_number(),
    cum_duration = cumsum(duration)
  )

ggplot(line_data, aes(x = video_index, y = cum_duration, colour = condition)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2, alpha = 0.6) +
  scale_colour_manual(values = pal2) +
  labs(
    title    = "Fig 3 · Line Plot: Cumulative Video Duration by Condition",
    subtitle = "Verifying that total encoding time is matched across conditions",
    x = "Video Index", y = "Cumulative Duration (seconds)", colour = "Condition"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))
```

---

### Fig 4 — Heat Map: Duration Bins by Condition

> **What it shows:** How videos are distributed across duration bins (e.g. 0–10s, 11–20s) for
> each condition. A well-matched stimulus set should show similar bin counts across both conditions.

```{r heatmap-duration}
movies %>%
  mutate(
    dur_bin = cut(duration,
                  breaks = c(0, 10, 20, 30, 40, 60, Inf),
                  labels = c("0-10s", "11-20s", "21-30s", "31-40s", "41-60s", "60s+"))
  ) %>%
  count(condition, dur_bin) %>%
  complete(condition, dur_bin, fill = list(n = 0)) %>%
  ggplot(aes(x = condition, y = dur_bin, fill = n)) +
  geom_tile(colour = "white", linewidth = 0.6) +
  geom_text(aes(label = ifelse(n > 0, n, "")), size = 4.5) +
  scale_fill_gradient(low = "#E3F2FD", high = "#1565C0") +
  labs(
    title    = "Fig 4 · Heat Map: Video Count by Duration Bin and Condition",
    subtitle = "Checking that clip length distributions are comparable",
    x = "Condition", y = "Duration Bin", fill = "Count"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))
```

---

### Fig 5 — Waffle Chart: Repeat vs Non-Repeat Videos per Condition

> **What it shows:** Within each condition, how many videos were repeats (vigilance check) vs
> unique clips. The `is_repeat` column is a methodological detail — repeated videos were used
> only to check participant attentiveness during encoding and are not part of the main analysis.

```{r waffle-repeat, fig.height=4}
movies %>%
  mutate(repeat_lbl = ifelse(is_repeat == 1, "Repeat (Vigilance)", "Non-Repeat (Stimulus)")) %>%
  arrange(condition, repeat_lbl) %>%
  group_by(condition) %>%
  mutate(
    i   = row_number(),
    row = ceiling(i / 5),
    col = (i - 1) %% 5 + 1
  ) %>%
  ggplot(aes(x = col, y = -row, fill = repeat_lbl)) +
  geom_tile(colour = "white", linewidth = 1, width = 0.9, height = 0.9) +
  facet_wrap(~condition) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title    = "Fig 5 · Waffle Chart: Stimulus Composition per Condition",
    subtitle = "Each square = 1 video; repeated videos used as vigilance check only",
    fill     = NULL
  ) +
  theme_void(base_size = 12) +
  theme(
    plot.title      = element_text(face = "bold", hjust = 0.5),
    strip.text      = element_text(face = "bold", size = 11),
    legend.position = "bottom"
  )
```

---

## Section 2 — Recognition Task Structure

*These plots describe the target and lure frames used in the recognition memory test.*

---

### Fig 6 — Bar Chart: Frame Type Distribution

> **What it shows:** How many target frames are Before-Boundary (BB) vs Event-Middle (EM).
> This is central to Research Questions 2 and 3 — BB frames test boundary-related memory
> encoding, while EM frames serve as a control for non-boundary memory.

```{r bar-frametypes}
tl %>%
  count(frame_type) %>%
  ggplot(aes(x = reorder(frame_type, -n), y = n, fill = frame_type)) +
  geom_col(colour = "white", width = 0.55) +
  geom_text(aes(label = n), vjust = -0.4, size = 4.5) +
  scale_fill_manual(values = pal_ft) +
  labs(
    title    = "Fig 6 · Bar Chart: Target Frame Types in Recognition Task",
    subtitle = "BB = Before-Boundary (key experimental frames); EM = Event-Middle (control frames)",
    x = "Frame Type", y = "Count"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"), legend.position = "none")
```

---

### Fig 7 — Pie Chart: BB vs EM Proportion

> **What it shows:** The proportion of BB vs EM frames across all recognition trials.
> An equal split would mean the task is balanced between boundary-sensitive and
> boundary-insensitive memory probes.

```{r pie-frametypes, fig.width=6, fig.height=5}
tl %>%
  count(frame_type) %>%
  mutate(
    pct   = n / sum(n),
    label = paste0(frame_type, "\n", percent(pct, 1))
  ) %>%
  ggplot(aes(x = "", y = pct, fill = frame_type)) +
  geom_col(width = 1, colour = "white") +
  coord_polar("y") +
  geom_text(aes(label = label), position = position_stack(vjust = 0.5), size = 4) +
  scale_fill_manual(values = pal_ft) +
  labs(
    title = "Fig 7 · Pie Chart: Proportion of Frame Types",
    fill  = NULL
  ) +
  theme_void(base_size = 12) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.position = "none")
```

---

## Section 3 — Participant Demographics

*These plots describe who participated in each condition. Key checks are whether the Natural (NB)
and Abrupt (AB) groups are comparable on age, gender, handedness and vision.*

---

### Fig 8 — Bar Chart: Gender by Condition

> **What it shows:** The gender breakdown within each condition group.
> The two groups should be roughly comparable to rule out gender as a confound.

```{r bar-gender}
demog %>%
  count(condition, Gender) %>%
  ggplot(aes(x = condition, y = n, fill = Gender)) +
  geom_col(position = "dodge", width = 0.6, colour = "white") +
  geom_text(aes(label = n), position = position_dodge(0.6), vjust = -0.4, size = 4) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title    = "Fig 8 · Bar Chart: Gender Distribution by Condition",
    subtitle = "Checking gender balance across Natural (NB) and Abrupt (AB) groups",
    x = "Condition", y = "Count", fill = "Gender"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))
```

---

### Fig 9 — Violin Plot: Age Distribution by Condition

> **What it shows:** The spread of participant ages in each condition.
> Age should be similarly distributed across NB and AB groups for the comparison to be valid.

```{r violin-age}
ggplot(demog, aes(x = condition, y = Age, fill = condition)) +
  geom_violin(trim = FALSE, alpha = 0.7, colour = "grey40") +
  geom_boxplot(width = 0.08, fill = "white", outlier.shape = NA) +
  geom_beeswarm(size = 2.5, alpha = 0.6, cex = 2) +
  scale_fill_manual(values = pal2) +
  labs(
    title    = "Fig 9 · Violin Plot: Age Distribution by Condition",
    subtitle = "Individual points overlaid to show each participant",
    x = "Condition", y = "Age (years)"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"), legend.position = "none")
```

---

### Fig 10 — Mosaic Chart: Vision × Condition

> **What it shows:** The cross-tabulation of vision type (Normal vs Corrected to Normal) across
> conditions. Tile area reflects cell frequency — a good balance means roughly equal proportions
> in both conditions.

```{r mosaic-vision}
demog %>%
  mutate(Vision = str_trim(Vision)) %>%
  count(condition, Vision) %>%
  ggplot(aes(x = Vision, y = n, fill = condition)) +
  geom_col(position = "dodge", width = 0.6, colour = "white") +
  geom_text(aes(label = n), position = position_dodge(0.6), vjust = -0.4, size = 4) +
  scale_fill_manual(values = pal2) +
  labs(
    title    = "Fig 10 · Bar Chart: Vision Type by Condition",
    subtitle = "Checking vision balance across Natural (NB) and Abrupt (AB) groups",
    x = "Vision Type", y = "Count", fill = "Condition"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title  = element_text(face = "bold"),
    axis.text.x = element_text(angle = 15, hjust = 1)
  )
```

---

### Fig 11 — Spider / Radar Chart: Demographic Summary by Condition

> **What it shows:** A multi-axis overview of key demographic counts per condition —
> N participants, mean age, N female, N right-handed, N normal vision.
> Useful for a quick at-a-glance balance check across all demographic variables simultaneously.

```{r radar-demog, fig.width=7, fig.height=6}
radar_stats <- demog %>%
  group_by(condition) %>%
  summarise(
    N_Participants = n(),
    Mean_Age       = mean(Age, na.rm = TRUE),
    N_Female       = sum(Gender == "Female", na.rm = TRUE),
    N_RightHanded  = sum(Handedness == "Right handed", na.rm = TRUE),
    N_NormalVision = sum(str_detect(Vision, "Normal"), na.rm = TRUE)
  )

radar_mat <- radar_stats %>%
  select(-condition) %>%
  as.data.frame()

rownames(radar_mat) <- radar_stats$condition

max_row <- apply(radar_mat, 2, max) * 1.15
min_row <- apply(radar_mat, 2, min) *  0

radar_df <- rbind(max_row, min_row, radar_mat)

radarchart(
  radar_df,
  axistype   = 1,
  pcol       = c("#4CAF50", "#E91E63"),
  pfcol      = scales::alpha(c("#4CAF50", "#E91E63"), 0.3),
  plwd       = 2,
  cglcol     = "grey70", cglty = 1,
  axislabcol = "grey40",
  vlcex      = 0.85,
  title      = "Fig 11 · Radar Chart: Demographic Summary by Condition"
)

legend(
  "topright",
  legend = rownames(radar_mat),
  col    = c("#4CAF50", "#E91E63"),
  lty = 1, lwd = 2, bty = "n", pch = 20
)
```

---

*End of initial visualizations. Add your observations and interpretations in the space above each figure.*
